Pairwise-complete correlation considered dangerous

### 2015-05-27 missing.Rmd
references

This note warns about potentially misleading results when using the
`use=pairwise.complete.obs` and related options in R's `cor`
and `cov` functions. The note illustrates potential issues using
a very simple pathological example and lists alternative approaches
to handling missing data and references about them.

@@

R includes facilities for handling missing values for all native data types. Perhaps counterintuitively, marking a value as missing conveys the information that the value is not known. Donald Rumsfeld might call it a "known unknown^[https://goo.gl/gGYcJe]." Upon encountering a missing value, we can deal with it by simply omitting it, imputing it somehow, or through several other possible approaches. R does a good job of making our choice of missing value approach explicit.

>>>>

This note warns about potentially misleading results when using the
`use=pairwise.complete.obs` and related options in R's `cor`
and `cov` functions. Pitfalls are illustrated using
a very simple pathological example followed by a brief list
of alternative ways to deal with missing data and some references about them.

@@

R includes excellent facilities for handling missing values for all native data types. Perhaps counterintuitively, marking a value as missing conveys the information that the value is not known. Donald Rumsfeld might call it a "known unknown^[https://goo.gl/gGYcJe]." Upon encountering a missing value, we can deal with it by simply omitting it, imputing it somehow, or through several other possible approaches. R does a good job of making our choice of missing value approach explicit.
